---
title: "Analysis on solo-stake"
author: "Jonas Gehrlein"
date: "2/17/2021"
output: html_document
---

```{r}
# Specify if code should be shown in export
show_code = FALSE 
# Specify which network you would like to analyze. Must be "polkadot" or "kusama".
chain = "polkadot"
# Specify which session you would like to analyze. Must be >= 419 for Polkadot and >= 5994 for KSM.
session = 1599
```

```{r, echo=FALSE}
#Setting up the denomination KSM and (new) DOT:
network_identifier = ifelse(chain == "polkadot", "DOT", "KSM")
if(chain=="polkadot"){
  normalization = 1/10000000000
} else {
  normalization = 1/1000000000000
}
```

```{r setup, fig.width=4, fig.height=3, warning=FALSE, message=FALSE, echo=FALSE}
#Load the required R packages:
library(plyr)
library(dplyr)
library(rmarkdown)
library(ggplot2)
library(tidyverse)
library(dbplyr)
library(igraph)
library(network)
library(tidygraph)
library(ggraph)
library(splitstackshape)
library(visNetwork)
library(networkD3)
library(htmlwidgets)
library(REAT)
```


```{r, echo=FALSE}
#Import the data into the data-frames:
nominators <- read.csv(url(paste("https://storage.googleapis.com/watcher-csv-exporter/", chain , "_nominators_session_", session, ".csv", sep=(""))))
validators <- read.csv(url(paste("https://storage.googleapis.com/watcher-csv-exporter/", chain , "_validators_session_", session, ".csv", sep=(""))))
```


```{r, echo=FALSE}
# Preparation
solo_stake_nominators <- subset(nominators, num_targets==1,
                  select=c(bonded_amount,targets))

solo_stake_nominators <- subset(solo_stake_nominators, select=c(bonded_amount, targets))

collapse <- solo_stake_nominators %>%
  group_by(targets) %>%
  summarize(solo_stake_sum = sum(bonded_amount))

collapse$ID = collapse$targets
validators$ID = validators$stash_address

merged <- merge(collapse, validators, by="ID", all.y=TRUE)
merged$solo_stake_sum <- ifelse(is.na(merged$solo_stake_sum), 0 , merged$solo_stake_sum)

validators_with_solo_stake = subset(merged, select=c(solo_stake_sum, name, stash_address, commission_percent, self_stake, total_stake))

# Adding the short identity
validators_with_solo_stake$has_identity = ifelse(validators_with_solo_stake$name == "",0,1)

# Only parse the first 4 characters of the identity string. Fortunately, validators follow a naming pattern where the first few characters are identical.
validators_with_solo_stake$short_identity = substr(validators_with_solo_stake$name, start = 1, stop = 4)

# Replace empty name with indicator that validator is pseudo-anonymous
validators_with_solo_stake$short_identity = ifelse(validators_with_solo_stake$short_identity == "","pseudo-a.",validators_with_solo_stake$short_identity)

tmp <- validators_with_solo_stake %>% 
    count(short_identity)
validators_with_solo_stake <- merge(validators_with_solo_stake,tmp, by="short_identity", all.x = TRUE)

# normalize all planck values to DOT/KSM
validators_with_solo_stake$solo_stake_sum = validators_with_solo_stake$solo_stake_sum * normalization
validators_with_solo_stake$self_stake = validators_with_solo_stake$self_stake * normalization
validators_with_solo_stake$total_stake = validators_with_solo_stake$total_stake * normalization
nominators$bonded_amount = nominators$bonded_amount * normalization
solo_stake_nominators$bonded_amount = solo_stake_nominators$bonded_amount*normalization
```
# Summary
In the following, several perspectives on solo-stake are taken. The main focus is on larger validator operators which impose the risk of spinning up node after node, transferring their self-stake away and imposing a nothing-at-stake risk. The fourth analysis (Solostake #4) indicates that some of the candidates for such an attack (i.e., many nodes, little commission and little self-stake) do have significant number of solo-stake and imposing a minimum might not work so much on them.  

# Distribution of solo-stake for each validator
Specifications: All individual nodes

This graph plots the sum of solo-stake for **every** validator currently in the network.

```{r, echo=FALSE}
solo_stake_sum = validators_with_solo_stake$solo_stake_sum
df_graph = as.data.frame(solo_stake_sum)
df_graph$validators = 1:nrow(df_graph)
df_graph$solo_stake_sum = sort(df_graph$solo_stake_sum, decreasing=FALSE)
ggplot(data=df_graph, aes(x=validators, y=solo_stake_sum))+ geom_bar(stat="identity")
``` 

# Solo-stake #1
Specifications: Grouped by operators - No pseudoanonymous nodes, operator-size > 4 nodes.

```{r, echo=FALSE}
# Collapse the data based on the short_identity taking the mean of self_stake and commission and the sum of solo_stake
# Super annoying, I want to take the mean and the sum of the solo_stake_sum therefore I first have to duplicate it.
validators_with_solo_stake$solo_stake_mean = validators_with_solo_stake$solo_stake_sum

df_table = merge(
      aggregate(cbind(solo_stake_sum) ~ short_identity, validators_with_solo_stake, sum), 
      aggregate(cbind(self_stake, commission_percent, n, solo_stake_mean) ~ short_identity, validators_with_solo_stake, mean)
            )
df_table$fraction <- round(df_table$n / sum(df_table$n),digits=3)

df_table = df_table %>% 
  rename(
    number_nodes = n,
    self_stake_mean = self_stake,
    commission_percent_mean = commission_percent
    )

# Order by number of nodes
df_table <- df_table[order(-df_table$number_nodes),]
rownames(df_table) <- NULL
# Table Output
knitr::kable(df_table, format="simple", format.args = list(scientific = FALSE))

#  Graph Output
df_graph = df_table$solo_stake_sum
df_graph = as.data.frame(df_graph)

df_graph = df_graph %>% 
  rename(
    solo_stake_sum = df_graph
    )
df_graph$validators = 1:nrow(df_graph)
df_graph$solo_stake_sum = sort(df_graph$solo_stake_sum, decreasing=FALSE)
ggplot(data=df_graph, aes(x=validators, y=solo_stake_sum))+ geom_bar(stat="identity")
```

# Solo-stake #2
Specifications: Individual nodes conditional on operators - No pseudoanonymous nodes, operator-size > 4 nodes.

```{r, echo=FALSE}
df_table = subset(validators_with_solo_stake, n > 4 & short_identity != "pseudo-a.")
df_table = subset(df_table, select=c(n, name, solo_stake_sum, self_stake, commission_percent))

# Order by number of nodes
df_table <- df_table[order(-df_table$n),]
rownames(df_table) <- NULL

# Table Output
knitr::kable(df_table, format="simple", format.args = list(scientific = FALSE))

# Graph Output
df_graph = df_table$solo_stake_sum
df_graph = as.data.frame(df_graph)

df_graph = df_graph %>% 
  rename(
    solo_stake_sum = df_graph
    )
df_graph$validators = 1:nrow(df_graph)
df_graph$solo_stake_sum = sort(df_graph$solo_stake_sum, decreasing=FALSE)
ggplot(data=df_graph, aes(x=validators, y=solo_stake_sum))+ geom_bar(stat="identity")
``` 

# Solo-stake #3
Specifications: Individual nodes conditional on operators - No pseudoanonymous nodes, operator-size > 4 nodes, commission_percent < 100%.

```{r, echo=FALSE}
df_table = subset(validators_with_solo_stake, n > 4 & short_identity != "pseudo-a." & commission_percent < 100)
df_table = subset(df_table, select=c(n, name, solo_stake_sum, self_stake, commission_percent))

# Order by number of nodes
df_table <- df_table[order(-df_table$n),]
rownames(df_table) <- NULL

# Table Output
knitr::kable(df_table, format="simple", format.args = list(scientific = FALSE))

# Graph Output
df_graph = df_table$solo_stake_sum
df_graph = as.data.frame(df_graph)

df_graph = df_graph %>% 
  rename(
    solo_stake_sum = df_graph
    )
df_graph$validators = 1:nrow(df_graph)
df_graph$solo_stake_sum = sort(df_graph$solo_stake_sum, decreasing=FALSE)
ggplot(data=df_graph, aes(x=validators, y=solo_stake_sum))+ geom_bar(stat="identity")
```

# Solo-stake #4
Specifications: Individual nodes conditional on operators - No pseudoanonymous nodes, operator-size > 4 nodes, commission_percent < 3%.

```{r, echo=FALSE}
df_table = subset(validators_with_solo_stake, n > 4 & short_identity != "pseudo-a." & commission_percent < 3)
df_table = subset(df_table, select=c(n, name, solo_stake_sum, self_stake, commission_percent))

# Order by number of nodes
df_table <- df_table[order(-df_table$n),]
rownames(df_table) <- NULL

# Table Output
knitr::kable(df_table, format="simple", format.args = list(scientific = FALSE))

# Graph Output
df_graph = df_table$solo_stake_sum
df_graph = as.data.frame(df_graph)

df_graph = df_graph %>% 
  rename(
    solo_stake_sum = df_graph
    )
df_graph$validators = 1:nrow(df_graph)
df_graph$solo_stake_sum = sort(df_graph$solo_stake_sum, decreasing=FALSE)
ggplot(data=df_graph, aes(x=validators, y=solo_stake_sum))+ geom_bar(stat="identity")
```

# Flowchart of Solo-stake
```{r, echo=FALSE}
# Sankey Network of Bondflow
# Define quantiles of bonded_amount to group nominators into those groups
nominators_sankey = subset.data.frame(nominators, num_targets == 1)
quantiles = quantile(nominators_sankey$bonded_amount, c(.10, .20, .30, .40, .50, .60, .70, .80, .90, .95, .98)) 

# Check in which quantile each nominator is
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount <= quantiles[1], " poorest 10%", "")
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount > quantiles[1] & nominators_sankey$bonded_amount <= quantiles[2], " between 10-20%", nominators_sankey$quantiles)
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount > quantiles[2] & nominators_sankey$bonded_amount <= quantiles[3], " between 20-30%", nominators_sankey$quantiles)
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount > quantiles[3] & nominators_sankey$bonded_amount <= quantiles[4], " between 30-40%", nominators_sankey$quantiles)
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount > quantiles[4] & nominators_sankey$bonded_amount <= quantiles[5], " between 40-50%", nominators_sankey$quantiles)
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount > quantiles[5] & nominators_sankey$bonded_amount <= quantiles[6], " between 50-60%", nominators_sankey$quantiles)
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount > quantiles[6] & nominators_sankey$bonded_amount <= quantiles[7], " between 60-70%", nominators_sankey$quantiles)
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount > quantiles[7] & nominators_sankey$bonded_amount <= quantiles[8], " between 70-80%", nominators_sankey$quantiles)
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount > quantiles[8] & nominators_sankey$bonded_amount <= quantiles[9], " between 80-90%", nominators_sankey$quantiles)
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount > quantiles[9] & nominators_sankey$bonded_amount <= quantiles[10], " between 90-95%", nominators_sankey$quantiles)
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount > quantiles[10] & nominators_sankey$bonded_amount <= quantiles[11], " between 95-98%", nominators_sankey$quantiles)
nominators_sankey$quantiles = ifelse(nominators_sankey$bonded_amount > quantiles[11], " richest 2%", nominators_sankey$quantiles)

#nominators = subset(nominators, quantiles != " richest 2%") # if we want to exclude the richest 2%
#validators = subset(validators, validators$commission_percent<100) # if we want to exclude the 100% validators

x=cSplit(nominators_sankey, c("targets"), ",", direction = "long")

y = as.data.frame(cbind(name = paste(validators$name), targets = paste(validators$stash_address)))

x <- merge(x, y, by="targets")

x$Weight = x$bonded_amount
x$Weight = as.numeric(x$Weight)
# Check if validator has identity
x$has_identity = ifelse(x$name == "",0,1)

# Only parse the first 4 characters of the identity string. Fortunately, validators follow a naming pattern where the first few characters are identical.
x$short_identity = substr(x$name, start = 1, stop = 4)

# Replace empty name with indicator that validator is pseudo-anonymous
x$short_identity = ifelse(x$short_identity == "","pseudo-a.",x$short_identity)

# Use newly generated identities (operators and pseudo-anonymous validators) as new destination indicator in network
x$Destination = x$short_identity

# Create the network as tibble object
network <- tibble(x)

network$Source = network$quantiles

# Cast objects to character type
network$Source = as.character(network$Source)
network$Destination = as.character(network$Destination)

# Generate sources data frame
sources <- network %>%
  distinct(Source) %>%
  dplyr::rename(label = Source)

# Generate destinations data frame
destinations <- network %>%
  distinct(Destination) %>%
  dplyr::rename(label = Destination)

# Create nodes 
nodes <- full_join(sources, destinations, by = "label")
nodes <- nodes %>% rowid_to_column("id")

# Create routes
per_route <- network %>%  
  group_by(Source, Destination) %>%
  ungroup()

# Create edges
edges <- per_route %>% 
  left_join(nodes, by = c("Source" = "label")) %>% 
  dplyr::rename(from = id)

edges <- edges %>% 
  left_join(nodes, by = c("Destination" = "label")) %>% 
  dplyr::rename(to = id)

edges <- select(edges, from, to, Weight)

nodes_d3 <- as.data.frame(mutate(nodes, id = id - 1))
edges_d3 <- as.data.frame(mutate(edges, from = from - 1, to = to - 1))

# Make sankeyNetwork object
sankey <- sankeyNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to", 
                        NodeID = "label", Value = "Weight", fontSize = 16, unit = "Bond", height=1300, width=1000, nodeWidth=80, iterations=50, margin=list("left"=100, "right"=100))

## Changes color to green
sankey_rendered <- htmlwidgets::onRender(sankey,
                                         'function(el, x) {
    d3.selectAll(".node text")
        .style("fill", "black");
  }'
)

## Changes the position of the leaf-nodes to right
sankey_rendered <- htmlwidgets::onRender(
  sankey_rendered,
  '
function(el,x){
  var node_text = d3.select(el)
    .selectAll(".node text")
    .attr("x", 6 + x.options.nodeWidth)
    .attr("text-anchor", "start");
}
'
)
## Change position of the parent nodes (starting with a space) to the left

sankey_rendered <- htmlwidgets::onRender(
  sankey_rendered,
  '
  function(el,x){
  // select all our node text
  d3.select(el)
  .selectAll(".node text")
  .filter(function(d) { return d.name.startsWith(" "); })
  .attr("x", x.options.nodeWidth - 16)
  .attr("text-anchor", "end");
  }
  '
)
sankey_rendered
``` 

